Bicycle Balance Assistance Reduces Probability of Falling at Low Speeds When Subjected to Mechanical Perturbations

### 2024-06-13 main.tex
Fill in some missing data

\section{Methods}
%
\subsection{Bicycle}
%
We modified a standard electric bicycle (Gazelle Grenoble C7\todo{Double check
this is the right model}, Dieren, The Netherlands) with a custom motor in the
steering assembly capable of applying up to 7~\si{\newton\meter} of torque
between the head tube and steer tube, see
Figure~\ref{fig:balance-assist-bicycle}.  A custom motor controller converts
the commanded torque to applied torque. We measure the speed of the rear wheel
with an encoder (Magura XXX) and measure the body fixed roll rate of the
bicycle with a MEMs rate gyroscope (Company, Location). The balance assist
control algorithm is implemented on a microprocessor (Teensy, Company,
Location) and data from all sensors is logged with a CAN bus (Company,
Location) at approximately X Hz.

@@

\subsection{Perturbations}
%
We apply external longitudinal forces to the ends of each handlebar using an
adapted Bump'Em system~\cite{Tan2020} which is designed with four motors
working in tandem, Figure~\ref{fig:setup-diagram}. The four motors are
programmed to apply a light force at all times to keep the ropes taught and to
track a commanded force profile using a PID controller running on a
microprocessor (Arduino XXX, Italy). We measure the force applied by each motor
at the handlebar via four inline load cells, each with a max load of
XX~\si{\newton}. The commanded force profiles are designed to apply an external
pulsive torque to the front assembly (handlebars, forks, wheel) at magnitudes
varying from X to X. The four motors are arranged at the four corners of a
1~\si{\meter} wide treadmill (make, Company, City) that can reach speeds of
18~\si{\kilo\meter\per\hour}. The general design is described in detail in the
Van De Velde's MSc thesis~\citep{vandeVelde2022a} and the physical arrangement
is shown in Figure~\ref{fig:participant-in-set-up}. Our modifications relative
to that described in the thesis include simplifying the controller with a
inexpensive microcontroller and the use of a simpler safety harness.
Figure~\ref{fig:perturbation_0} shows an example perturbation.
%
% NOTE : I get a compile error with the \protect calls here, not sure why.
\begin{figure}
  \centering
  \subcaptionbox{A participant on the bicycle in the safety harness with the
    Bump'Em motor lines attached to the ends of the
    handlebars.\protect\label{fig:participant-in-set-up}}
    {\includegraphics[width=75mm]{figures/participant-in-set-up.png}}
  \subcaptionbox{Top view diagram of bicycle handlebar perturbation
    system.\protect\label{fig:setup-diagram}}
    {\includegraphics[width=65mm]{figures/setup-diagram.png}}
  \caption{Diagrams showing the Bump'Em system arranged to apply handlebar
    perturbations.}
  \label{fig:experimental-setup}
\end{figure}

@@

The subjects were divided into two groups. The first group performed the
protocol at a belt speed of 2.8~\si{\meter\per\second}
(10~\si{\kilo\meter\per\hour}) with the controller gain at $g=8$ and the second
group performed the protocol at a belt speed of 1.7~\si{\meter\per\second}
(6~\si{\kilo\meter\per\hour}). The 10~\si{\kph} group's experiments occurred X
weeks before the 6~\si{\kilo\meter\per\hour} group.

Subjects wore a helmet and fall safety harness attached to the ceiling
Figure~\ref{fig:participant-in-set-up}. We allowed them to practice riding on
the treadmill until they indicated they were comfortable enough to have
perturbations applied. For most, this was less than a 10~\si{\minute} warm up.
We then asked the rider to ride for 90~\si{\second}, attempting to maintain the
location of their front wheel on the center line of the treadmill. We define a
``fall'' on the treadmill by two criteria: the rider removes their foot from
the pedal and places it on the ground or the bicycle wheel exceeds the width of
the treadmill. We then applied perturbations in random directions (clockwise or
counter clockwise) increasing the magnitude by X~\si{\newton\meter} each time.
Figure~\ref{fig:perturbation-sequence} shows an example result of a
perturbation.  We log the magnitude that causes the first fall to characterize
that subject's nominal fall threshold.  Following this we apply 20
perturbations of random magnitudes to the cyclist while they ride at a constant
speed and record which perturbations cause a fall. We let the cyclist rest and
then perform another 20 perturbations. We randomize whether the balance assist
system is on or off during the first or second set of perturbations.
%
\begin{figure}
  \centering
  \subcaptionbox{Start}{\includegraphics[width=40mm]{figures/start.png}}
  \subcaptionbox{During}{\includegraphics[width=40mm]{figures/during.png}}
  \subcaptionbox{After}{\includegraphics[width=40mm]{figures/after.png}}
  \subcaptionbox{Recovery}{\includegraphics[width=40mm]{figures/recovery.png}}
  \caption{Video frames depicting the application of a perturbation and the
  rider's response and recovery.}
  \label{fig:perturbation-sequence}
\end{figure}

>>>>

\section{Methods}
%
\subsection{Bicycle}
%
We modified a standard electric bicycle (Gazelle Grenoble C7\todo{Double check
this is the right model}, Dieren, The Netherlands) with a custom motor in the
steering assembly capable of applying up to 7~\si{\newton\meter} of torque
between the head tube and steer tube, see
Figure~\ref{fig:balance-assist-bicycle}.  A custom motor controller converts
the commanded torque to applied torque. We measure the speed of the rear wheel
with an encoder (Magura XXX) and measure the body fixed roll rate of the
bicycle with a MEMs rate gyroscope (CRS 03-02s, Silicon Sensing, United Kingdom).
\todo{Do you mean the gyro inside the bicycle or the external? Currently filled in the 
data of the external one. Not sure what make and model is inside the bicycle}
The balance assist control algorithm is implemented on a microprocessor (Teensy, PJRC,
USA) and data from all sensors is logged with a CAN bus (CanEdge2, CSS Electronics,
Denmark) at approximately 100 Hz. \todo{Sample rate is not the same for each
sensor: wheel speed, imu = 1ms, steer torque command is updated every 10ms, 
can log is written every 10ms}

@@

\subsection{Perturbations}
%
We apply external longitudinal forces to the ends of each handlebar using an
adapted Bump'Em system~\cite{Tan2020} which is designed with four motors
working in tandem, Figure~\ref{fig:setup-diagram}. The four motors are
programmed to apply a light force at all times to keep the ropes taught and to
track a commanded force profile using a PID controller running on a
microprocessor (Arduino Mega 2560, Italy). We measure the force applied by each 
motor at the handlebar via four inline load cells, each with a max load of
250~\si{\newton}. One load cell was damaged during pilot experiments, and is
replaced by a load cell with a max load of 500~\si{\newton}. The commanded 
force profiles are designed to apply an external pulsive torque to the front 
assembly (handlebars, forks, wheel) at magnitudes varying from 20 to 200. The four
motors are arranged at the four corners of a 1~\si{\meter} wide treadmill 
(EC-90 Flat, Maxon Group, Switzerland) that can reach speeds of 18~\si{\kilo\meter\per\hour}.
The general design is described in detail in the Van De Velde's MSc 
hesis~\citep{vandeVelde2022a} and the physical arrangement  is shown in 
Figure~\ref{fig:participant-in-set-up}. Our modifications relative to that 
described in the thesis include simplifying the controller with a inexpensive 
microcontroller and the use of a simpler safety harness. 
Figure~\ref{fig:perturbation_0} shows an example perturbation.
%
% NOTE : I get a compile error with the \protect calls here, not sure why.
\begin{figure}
  \centering
  \subcaptionbox{A participant on the bicycle in the safety harness with the
    Bump'Em motor lines attached to the ends of the
    handlebars.\protect\label{fig:participant-in-set-up}}
    {\includegraphics[width=75mm]{figures/participant-in-set-up.png}}
  \subcaptionbox{Top view diagram of bicycle handlebar perturbation
    system.\protect\label{fig:setup-diagram}}
    {\includegraphics[width=65mm]{figures/setup-diagram.png}}
  \caption{Diagrams showing the Bump'Em system arranged to apply handlebar
    perturbations.}
  \label{fig:experimental-setup}
\end{figure}

@@

The subjects were divided into two groups. The first group performed the
protocol at a belt speed of 2.8~\si{\meter\per\second}
(10~\si{\kilo\meter\per\hour}) with the controller gain at $g=8$ and the second
group performed the protocol at a belt speed of 1.7~\si{\meter\per\second}
(6~\si{\kilo\meter\per\hour}). The 10~\si{\kph} group's experiments occurred 2
weeks before the 6~\si{\kilo\meter\per\hour} group.

Subjects wore a helmet and fall safety harness attached to the ceiling
Figure~\ref{fig:participant-in-set-up}. We allowed them to practice riding on
the treadmill until they indicated they were comfortable enough to have
perturbations applied. For most, this was less than a 10~\si{\minute} warm up.
We then asked the rider to ride for 90~\si{\second}, attempting to maintain the
location of their front wheel on the center line of the treadmill. \todo{The results 
from this are not included because the balance-assist was turned on for some participants}
We define a
``fall'' on the treadmill by two criteria: the rider removes their foot from
the pedal and places it on the ground or the bicycle wheel exceeds the width of
the treadmill. We then applied perturbations in random directions (clockwise or
counter clockwise), starting at 20~\si{\newton} and increasing the magnitude by 
20~\si{\newton\meter} until the participants falls.
Figure~\ref{fig:perturbation-sequence} shows an example result of a
perturbation.  We log the magnitude that causes the first fall to characterize
that subject's nominal fall threshold. Following this we apply 20
perturbations of random magnitudes to the cyclist while they ride at a constant
speed and record which perturbations cause a fall. We let the cyclist rest and
then perform another 20 perturbations. We randomize whether the balance assist
system is on or off during the first or second set of perturbations.
%
\begin{figure}
  \centering
  \subcaptionbox{Start}{\includegraphics[width=40mm]{figures/start.png}}
  \subcaptionbox{During}{\includegraphics[width=40mm]{figures/during.png}}
  \subcaptionbox{After}{\includegraphics[width=40mm]{figures/after.png}}
  \subcaptionbox{Recovery}{\includegraphics[width=40mm]{figures/recovery.png}}
  \caption{Video frames depicting the application of a perturbation and the
  rider's response and recovery.}
  \label{fig:perturbation-sequence}
\end{figure}
